---
title: "Intrinsics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

The following is in response to our 2nd week's assignment. Using the provided list of functions I will define each and demonstrate the function in use. 

#General Bits
1. help(topic)  -- opens the help menu to the R Documentation on topic written in the parantheses 
```{r}
help(plot)
```

2. ?topic -- similar to help(topic), opens R Documentation
```{r}
?distribution
```

3. ls() -- shows defined variables
```{r}
ls()
```

4. dir() -- shows files in directory
```{r}
dir()
```

5. list.files() -- lists the files in a directory/folder
```{r}
list.files()
```

#Input and Output
1. save() -- function to save data created in R in a specific file
```{r}
red<- seq(1,20,2)
save(red,file="seq.rmd")

```

2. load() -- reloads objects or datasets that were configured with               save()
```{r}
load("seq.rmd")
```

3. data() -- lists the data sets currently known or saved, loads specificed data sets
```{r}
data("iris")
```

4. library() -- lists the packages in R's library and can load specified packages in library
```{r}
library(ggplot2)
```

5. read.table() -- opens a document or file as a table
```{r} 
read.table(file="int_prac.rtf",sep=",")
```
 -- did not execute properly, don't know what wrong.

6. read.csv() -- function to read a csv file and convert to R 
```{r}
read.csv(file="iris.csv", header=TRUE, sep="")
```

7. scan() -- scans a file (??)
```{r}
scan(file = "iris.csv", what = list(""))
```

8. print() -- prints the content of specificed variable onto   screen  
```{r}
ans<- 5*8
print(ans)
```

9. cat() -- concatenates two separate strings 
```{r}
AL <-cat("ana",".","lakshin")
```

10. write.table() -- exports R data as a readable and downloadable file
```{r}
write.table(iris,file="iris.csv",row.names = TRUE)
```


#Data Creation
1. c() -- combines a vector so that a single variable could be assigned a set of numbers
```{r}
abs<- c(2,6,9,12)
```

2. from:to (where from and to are replaced with numbers, e.g. 1:10)     -- create a sequence of numbers
```{r}
bc <- 12:4
cd <- bc + 2
2:5 *2
```

3. seq() -- function to create a sequence of numbers
```{r}
sam<- seq(4,18,.25)
print(sam)
```

4. rep() -- replicates values in the specified variable
```{r}
rep(6:9, times=2)
```

5. data.frame() -- converts matrices into data frames, creates data frames
```{r}
yan <- matrix(seq(1,24,1), nrow=6, ncol = 4)
data.frame(yan)
```

6. list() -- creates a list
```{r}
kin<- list("a"=2.5, "b"= TRUE, "c"= 2:5)

#howdo you do this with names?

```

7. matrix() -- creates a martix with specified rows and columns
```{r}
yan <- matrix(seq(1,24,1), nrow=6, ncol = 4)
yan2 <- matrix(seq(25,50,1), nrow=6, ncol = 4)
```

8. factor() -- converts vector into factor, finds unique characters/strings & sorts alphabetically
```{r}
STEM<- c("science","math", "technology", "engineering","math")
STEM_factor <- factor(STEM)
print(STEM_factor)

```

9. rbind() -- combines two separate data matrices by row, has to be of same dataset or else combination meaningless
```{r}
rbind(yan,yan2)
```

10. cbind() -- combines two separate data matrices by column, has to be of same dataset or else combination meaningless
```{r}
cbind(yan, yan2)
```


#Slicing and Extracting Data
##Indexing vectors
1. x[n] nth element -- indexes the number or character by its position within the vector or matrix as specified by number in bracket
```{r}
g <- matrix(1:20, nrow = 2, ncol = 10)
g[2]
h <- c(4,8,21,34,51,83)
h [4]
```

2. x[-n] all but nth element -- indexes entire vector or matrix except the specified element
```{r}
g[-7]
```

3. x[1:n] first n elements -- indexes the sequence of numbers specified within the brackets from a vector or matrix
```{r}
g[2:4]
h[1:3]
```

4. x[-(1:n)] elements from n+1 to the end -- indexes all elements of a vector or matrix except those specified within the parentheses.
```{r}
h[-(2:4)]
g[-(5:10)]
```

5. x[c(1,4,2)] specific elements -- indexes specific (combinations of) numbers from a vector or matrix
```{r}
h[c(1,5)]
g[c(2,6,20)]
```

6. x[“name”] elements named “name”
```{r}
j <- c("john", "emily", "ana", "frank", "cam")
j["ana"]
class(j)

beb<- table(j)
b<- rep(beb,5)  # ???? not replicating?!
names(b)

b["john"]
# wont work as a vector. make it a table first
#or names function
```

7. x[x>3] all elements greater than 3 -- indexes numbers either greater than or less than the specified number
```{r}
g[g>12]
h[h<50]
```

8. x[x > 3 & x < 5] all elements between 3 and 5 -- indexes numbers with logical statements
```{r}
g[g > 6 & g< 12]
h[h > 20 & h < 40]
```

9. x[x %in% c(“a”,“and”,“the”)] all elements in given set
```{r}
"ana" %in% j

"john" %in% b == FALSE

```


## Indexing lists
```{r}
song<- list("Hotel California", 378, 4)

names(song) <- c("title", "duration", "track")

similar_song <- list (title = "Misery Business", duration= 190, track = 5)

song<- list (title = "Hotel California", duration = 378, track = 4, similar = similar_song)

str(song)

#https://www.youtube.com/watch?v=Px9VNWHja4M
```

1. x[n] list with elements n -- indexes main list's nth item
```{r}
song[2]
```

2. x[[n]] nt h element of the list 
```{r}
song[[4]][1]
#2nd song
```

3. x[[“name”]] element of the list named “name”
```{r}
song[["similar"]]["duration"]
```

4. x$name id.
```{r}
song$similar$duration

#how do i get to the 2nd "song" 
```


## Indexing matrices
```{r}
as<- c(31,123,67,1123,987)
ss<- c(122,56,897,32,45)
sb<- list(as,ss)
sbb<-matrix(data=as & ss, nrow=5, ncol = 2, byrow=FALSE, dimnames = NULL)

#why doesnt a list work in a matrix if they're both numeric
```

1. x[i,j] element at row i, column j
```{r}
sbb[1,2]
```

2. x[i,] row i
```{r}
sbb[4]
```

3. x[,j] column j
```{r}
sbb[,2]
```

4. x[,c(1,3)] columns 1 and 3
```{r}
sbb[c(2,3)]
```

5. x[“name”,] row named “name”
```{r}
first<-matrix(1:9, nrow=3)
rownames(first)<-(c("a","b","c"))

first["c",]
```


## Indexing data frames (matrix indexing plus the following)
```{r}
name<- c("ana","pete", "dave")
age<- c(22,32,45)
pass<-c(TRUE,FALSE,FALSE)

nap<-data.frame(name,age,pass)
nap
```

1. x[[“name”]] column named “name” -- indexes variables within specified column
```{r}
nap[["age"]]
```

2. x$nameid. -- indexes variables within specified column
```{r}
nap$pass
```


# Variable Conversion 
1. as.data.frame(x) -- converts r objects into data frame format
```{r}
as.data.frame(name)
```

2. as.numeric(x) -- change the composition of a vector into       numbers (if possible)
```{r}
dod <- c(5,"4",7)
as.numeric(dod)
```

3. as.logical(x) -- converts all non-zero numbers to True and all zeros as false
```{r}
ll<- c(0,4,7,2,0,1)
as.logical(ll)
```

4. as.character(x) -- converts numbers into characters
```{r}
cob<- 5.692
cob<- as.character(cob)
```


#Variable Information 
1. is.na(x) -- determines if any values are missing or not available (??)
```{r}
na <- c(4,7,NA,21)
is.na(na)
```

2. is.null(x) -- checks if there is a null variable
```{r}
is.null(age)
```

3. is.data.frame(x) -- checks if object is in a data frame format
```{r}
is.data.frame(nap)
```


4. is.numeric(x) -- checks if variable/object is a number (returns logical)
```{r}
is.numeric(cob)
```

5. is.character(x) -- checks if variable/item is a character class
```{r}
is.character(cob)
```

6. length(x) -- returns length of vector
```{r}
length(a)
```

7. dim(x) -- returns dimension of object(vector/list/matrix etc)
```{r}
dim(nap)
```

8. dimnames(x) -- lists dimensions/columns and their names
```{r}
dimnames(nap)
```

9. nrow(x) -- returns how many rows are in a matrix/dataframe & sets number of rows 
```{r}
nrow(nap)
```

10. ncol(x) -- returns number of columns in a matrix/dataframe & sets number of columns
```{r}
ncol(nap)
```

11. class() -- states a variable's class i.e. char, num, integer
```{r}
class(ans)
```

12. attributes() -- displays attributs of object
```{r}
attributes(yan)
```


#Data Selection and Manipulation
1. which.max() -- returns position of the highest value
```{r}
which.max(a)
```

2. which.min() -- returns position of the lowest value
```{r}
which.min(a)
```

3. which() -- returns position of which numbers fit whatever criteria you provide
```{r}
v <- c(6,3,7,1,8)
which(v %% 2 ==0)
```

4. sort() -- sorts a vector from lowest to highest
```{r}
sort(v)
```

5. unique() -- returns the unique values within a vector/list/matrix
```{r}
w <- c(1,1,3,6,6,3,4,4)
unique(w)
```

6. table() -- creates a table that keeps a count of the frequency of unique characters 
```{r}
vef<-c("sd","asf","ag","sd","rg","asf")
table(vef)
```

7. sample() -- takes a sample of numbers, random numbers under specified restrictions. 
```{r}
sample(100, 10, replace=FALSE)
```


#Math
1. max() -- returns the maximum number of a set or vector
```{r}
felix <- seq(3,56,1.2)
max(felix)
```

2. min() -- returns the minimum number of a set or vector
```{r}
min(felix)
```

3. range() -- returns the minimum and maximum of a set or vector
```{r}
range(felix)
```

4. sum() -- calculates the sum of a number set or vector 
```{r}
s_u <-sum(felix)
s_u
```

5. mean() -- function to find the mean of a number set or vector
```{r}
m_ <-mean(c(3,7,12,99,23,54))
m_
```

6. median() -- finds the median of a number set or vector
```{r}
nad<- seq(1,50,3)
z<- median(nad)
z
```


7. var() -- computes variance 
```{r}
ka <- c(1,5,3)
an <- c(4,5,8)

var(ka)
```

8. sd() -- computes standard deviation 
```{r}
sd(an)
```

9. cor() -- computes correlation 
```{r}
cor(ka,an)
```

10. round() -- rounds the value of input
```{r}
round(9/2)
```

11. abs() -- returns absolute value
```{r}
abs(-5)
```


#Matrices
```{r}
gre<-matrix(2:20, nrow=5)
```

1. t() -- transposes matrix
```{r}
t(first)
```

2. diag() -- returns numbers within the matrix that are in diagonal positions
```{r}
diag(first)
```

3. rowSums() -- returns sums of the rows
```{r}
rowSums(gre)
```

4. colSums() -- returns sums of the columns
```{r}
colSums(gre)
```

5. rowMeans() -- returns means of the rows
```{r}
rowMeans(gre)
```

6. colMeans() -- returns means of the columns
```{r}
colMeans(gre)
```


#Advanced Data processing
1. apply() -- applies specified functions on matrices. 2nd number is margin 1=rows, 2=columns
```{r}
gre
apply(gre,2,sum)
```

2. aggregate() --?
```{r, eval = FALSE}
aggregate(gre)
#dont understand the function
```


#Strings
1. paste() -- similar to concatenate, has separators
```{r}
rat1 <- "17R1"
paste(rat1, "box", sep = " is in ")
```

2. strsplit() -- 
```{r}

```

3. tolower() -- 
```{r}
alpha<- LETTERS
tolower(alpha)
```

4. toupper -- 
```{r}
alphab<- letters
toupper(alphab)
```


#Plotting
1. hist() - function for creating a histogram
```{r}
uk <- rnorm(50,mean=20,sd=1)
hist(uk)
```

2. plot() -- function for plotting a graph
```{r}
plot(rnorm(100,mean=38,sd=2))
```


#Distributions
1. rnorm()
```{r}
q<- rnorm(20, mean = 5, sd = 2)
plot(q)
```

2. runif() -- generates random numbers (how many, min, max)
```{r}
runif(10,5,50)
```


#Programming
1. Show that you can define a function
```{r}
mult4<- function(m)
  return(m*4)

mult4(10)
```

2. Show that you can write a for loop
```{r}
d<- c(81,21,52,36,19,22)
for(s in 1:4)
  d[s]<-d[s]+1
print(d)
```

3. Show that you can write a while loop
```{r}
dud<- c("hello","R")
ded<- 4:12
while(ded>10){
    print(dud)
    ded=ded+1
}

```

4. Show that you can write an if else statement
```{r}
a <- c(6,1,4,9)
ifelse(a%%2 == 0, "even","odd")
```

5. Explain what return() does inside a function, show you can use it
Return is used within functions to assign a number, value, r object to be brought back after the function completes. 
```{r}
add2<-function(x){
    return(x+2)
}
add2(5)
``` 

6. Explain what break() does, show you can use it
Break is a function that controls the flow of a loop, it stops the iterations and flows out. Inside nested loops the flow exists the innermost loop.
```{r}
f<- c(3,1,5,32,11,42)
f_num<-1:length(f)
for(g in f_num){
  if(f[g]<30){
    f[g]<- f[g]+3
}
  if(f[g]>30){
    break
    
  }
  }

print(f[g])
```