---
title: "Intrinsics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

The following is in response to our 2nd week's assignment. Using the provided list of functions I will define each and demonstrate the function in use. 

#General Bits
1. help(topic)  -- opens the help menu to the R Documentation on topic written in the parantheses 
```{r}
help(plot)
```

2. ?topic -- similar to help(topic), opens R Documentation
```{r}
?distribution
```

3. ls() -- shows defined variables
```{r}
ls()
```

4. dir() -- shows files in directory
```{r}
dir()
```

5. list.files() -- lists the files in a directory/folder
```{r}
list.files()
```

#Input and Output
1. save()
```{r}
red<- seq(1,20,2)
save(red,file="seq.rmd")

```

2. load() -- reloads objects or datasets that were configured with               save()
```{r}
load("seq.rmd")
```

3. data() -- 
```{r}
?data
```

4. library() --
```{r}

```

5. read.table() -- 
```{r}

```

6. read.csv() -- 
```{r}

```

7. scan() -- 
```{r}

```

8. print() -- 
```{r}

```

9. cat() -- concatenates two separate strings 
```{r}
AL <-cat("ana",".","lakshin")
```

10. write.table() -- 
```{r}
?write.table
```


#Data Creation
1. c()
2. from:to (where from and to are replaced with numbers, e.g. 1:10)
3. seq()
4. rep()
5. data.frame()
6. list()
7. matrix()
8. factor()
9. rbind()
10. cbind()

#Slicing and Extracting Data
##Indexing vectors
1. x[n] nth element
2. x[-n] all but nth element
3. x[1:n] first n elements
4. x[-(1:n)] elements from n+1 to the end
5. x[c(1,4,2)] specific elements
6. x[“name”] elements named “name”
7. x[x>3] all elements greater than 3
8. x[x > 3 & x < 5] all elements between 3 and 5
9. x[x %in% c(“a”,“and”,“the”)] all elements in given set

## Indexing lists
1. x[n] list with elements n
2. x[[n]] nt h element of the list
3. x[[“name”]] element of the list named “name”
4. x$name id.

## Indexing matrices
1. x[i,j] element at row i, column j
2. x[i,] row i
3. x[,j] column j
4. x[,c(1,3)] columns 1 and 3
5. x[“name”,] row named “name”

## Indexing data frames (matrix indexing plus the following)
1. x[[“name”]] column named “name” -- 
```{r}

```

2. x$nameid. -- 
```{r}

```


# Variable Conversion 
1. as.data.frame(x) -- 
```{r}

```

2. as.numeric(x) -- change the composition of a vector into       numbers (if possible)
```{r}
dod <- c(5,"4",7)
as.numeric(dod)
```

3. as.logical(x) -- 
```{r}

```

4. as.character(x) -- 
```{r}

```


#Variable Information 
1. is.na(x)
2. is.null(x)
3. is.data.frame(x)
4. is.numeric(x)
5. is.character(x)
6. length(x)
7. dim(x)
8. dimnames(x)
9. nrow(x)
10. ncol(x)
11. class()
12. attributes()

#Data Selection and Manipulation
1. which.max()
2. which.min()
3. which()
4. sort()
5. unique()
6. table()
7. sample()

#Math
1. max() -- returns the maximum number of a set or vector
```{r}
felix <- seq(3,56,1.2)
max(felix)
```

2. min() -- returns the minimum number of a set or vector
```{r}
min(felix)
```

3. range() -- returns the minimum and maximum of a set or vector
```{r}
range(felix)
```

4. sum() -- calculates the sum of a number set or vector 
```{r}
s_u <-sum(felix)
s_u
```

5. mean() -- function to find the mean of a number set or vector
```{r}
m_ <-mean(c(3,7,12,99,23,54))
m_
```

6. median() -- finds the median of a number set or vector
```{r}
nad<- seq(1,50,3)
z<- median(nad)
z
```

7. var() -- computes variance 
```{r}
?var
```

8. sd() -- computes standard deviation 
```{r}

```

9. cor() -- computes correlation 
```{r}

```

10. round() --
```{r}

```

11. abs() --
```{r}

```


#Matrices
1. t() -- 
```{r}

```

2. diag() -- 
```{r}

```

3. rowSums() -- 
```{r}

```

4. colSums() --
```{r}

```

5. rowMeans() --
```{r}

```

6. colMeans() --
```{r}

```


#Advanced Data processing
1. apply() -- 
```{r}

```

2. aggregate() --
```{r}

```


#Strings
1. paste() -- 
```{r}

```

2. strsplit() -- 
```{r}

```

3. tolower()
4. toupper

#Plotting
1. hist()
```{r}

```

2. plot()
```{r}
plot(rnorm(100,mean=38,sd=2))
```


#Distributions
1. rnorm()
2. runif()

#Programming
1. Show that you can define a function
2. Show that you can write a for loop
3. Show that you can write a while loop
4. Show that you can write an if else statement
5. Explain what return() does inside a function, show you can use it
6. Explain what break() does, show you can use it





